---
title: "CalendarHeatVisualization"
author: "Vinay Mandal"
date: "July 25, 2015"
output: html_document
---

I will be following tutorial from flowing data to generate "Calendar Heatmaps" for daily percent change in AAPL stock since 1981. I downloaded the dataset for this visualization from [Quandl](https://www.quandl.com/help/api).

```{r}
df <- read.csv('data/WIKI-AAPL.csv', header = T)
```

First few rows of the data look like this:

```{r, echo=FALSE}
head(df)
```

Now we are going to calculate the daily percentage changes in the stock price. We will use the opening and closing price of the stock to calculate this.

```{r}
df['Change_pct'] <- (df['Open'] - df['Close'])*100/df['Close']
df$date = as.Date(df$Date, "%Y-%m-%d")
head(df)
```

Since AAPL stock was publicly listed on 1980-12-12, we will pad the earlier dates in 1981 with "no change" (0) value. We do the same for rest of the 2015 dates.

```{r}
library(lubridate)
mindate <- as.Date(sprintf("%s-01-01", as.character(year(min(df$date)))))
maxdate <- as.Date(sprintf("%s-12-31", as.character(year(max(df$date)))))
allDates <- data.frame(date=seq.Date(mindate, maxdate, by="day"))
dfnew <- merge(df, allDates, by='date', all=TRUE)
dfnew['Change_pct'][is.na(dfnew['Change_pct'])] <- 0
```
Next, we are going to plot the calendar. First we will source the calendar code by Nathan Yau. I have made some minor modifications to it - nothing too fancy though. I have also created a function which plots a single year heatmap. We will use sapply to iterate over years and apply this function to plot the heatmap
```{r}
source('./calendarCustom.R')

plotSingle <- function(yr, df) {
  dfsubset <- subset(df, year(date) == yr)
  calendarFlow(dfsubset$date, dfsubset$Change_pct, palette="redgreen", main=toString(yr))
}

years <- seq(1980, 2015, 1)
par(mfrow = c(12, 3), oma=c(1, 1, 0, 1), mar=c(0, 0, 0, 0))
invisible(sapply(years, plotSingle, df=dfnew))
```